---
title: "ESM 206 Lab 5"
author: "Zoe Duerksen-Salm"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with 'lubridate' package
- Get counts of observations with count( )
- Uncount( ) rows
- One and two-sample t-test
- Create a heatmap with geom_tile( )

```{r, include = FALSE}

# include = FALSE > code will not show up*

# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  janitor::clean_names()

```

Use tidyr::uncount( ) to convert our lobster data from frequency format to case format

```{r}

# uncount() duplicates rows according to a weighting varaible (given by column lobster_count)
## uncount() makes it easy to get from frequency back to raw data

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# width = reduces the width of the geom scatter

# with a lot of observations, an easier way to look at distributions is using histograms and/or quantile quantile plots

## HISTORGRAM
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site),
                 show.legend = FALSE) +
  facet_wrap(~site, scales = "free")

# [ sclaes = "free" ] makes it so that the sale of each individual facet_wrap graph is on it's own scale dependnet on its own values. This can look a lot better, but can also make really misleading data (can be bad)

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)

```

#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse motn & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Now let's parse year and month using lubridate::month( ) and lubridate::year( )

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

```

Now let's find counts of observed lobsters based on different hierarchical groupings:

- Use the count( ) function - does a number of things all in one place
  - groups things based on how you tell it to group things
  - reports counts
  - then ungroups data


```{r}

# First... Let's count lobsters by year and month:

lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

# Then... Count lobsters JUST by year:

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_y

# Finally... Count lobsters ONLY by site:

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site

```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by( ) + n( )

- Using n() requires group by and summary
- Using tally only requires group by (and assumes you want to summarize)

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary

```

#### Find confidence intervals

If we calculate a 95% confidence interval based on a sample -- that means that if we took a bunch more samples, then in 95% of the intervals recorded, we would expect to find the population mean.

Use t.test( ) function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (... this will be covered conceptually in lectures Week 6)

```{r}

ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

## To pull a single column from a df to have all values show up as a vector you can use pull( ) function.
## This makes the df no longer a df, it is now a vector/matrix

t.test(ivee_lobster)

```

#### Two-sample t-test to compare means

If 2 samples are drawn from the same population - and they are perfect random samples- what would the difference in the means of the samples? It should be equal to zero.

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our ncessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}

# Method 1:

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest

```

There is a significant difference in lobster lenths between Naples and Mohawk Reef (t(df) = statistic, p< 0.001, alpha - 0.05).

The statement of significance should be the least interesting thing about your data. 

---

Different approach: using "model format"

```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

# ~ = as a function of/in reference to...

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

# You get the exact same results!! It is very unlikely that you would get variables this distinct if they were drawn from populations with the same mean.

# Reporting values/results from your t-test...

mn_ttest2$p.value
mn_ttest2$statistic

```

#### Calling values from your code chunks
I.e., referencing values/results from your analyses
Aka, in-line referencing

Here is the p-value: `r mn_ttest2$p.val`

Here is the t-statistic: `r mn_ttest2$statistic`

#### Now: a heatmap!

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

